---
title: "FIT2086 Assignment 3"
author: "Ian Wong 30612616"
date: "04/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

# Loading Libraries

```{r}
library(tidyverse)
```

# Question 1

## 1.

```{r}
concrete <- read_csv("concrete.ass3.2021.csv")
```

```{r}
fit <- lm(Strength ~ ., data=concrete)
summary(fit)
```
The fitted linear model is:

$$\begin{aligned}
\widehat{Strength}_i=-&3.02+0.12Cement_i+0.10Blast.Furnace.Slag_i+0.07Fly.Ash_i-0.19Water_i \\
&+0.42Superplasticizer_i+0.01Coarse.Aggregate_i+0.02Fine.Aggregate_i+0.13Age_i
\end{aligned}$$

The predictors that are possibly associated with the compressive strength of concrete are **Cement, Blast.Furnace.Slag, Fly.Ash, Water, Superplasticizer**, and **Age**. 
This is because compressive strength is most likely determined by the amount of each component substance within the concrete mixture.  

The three variables that appear to be the strongest predictors of compressive strength appear to be:  

  - **Age,**
  - **Cement,** and
  - **Blast Furnace Slag**

This is based on the p-values of these three predictors being the smallest compared to the other predictors according to the information gathered from `summary()`.

## 2.

If we were to use the Bonferroni procedure with $\alpha = 0.05$ and $p = 8$, we would instead reject the null hypothesis only if:
$$\begin{aligned}
\text{p-value } &< \frac{\alpha}{p} \\
&< \frac{0.05}{8} \\
&< 0.00625
\end{aligned}$$

Therefore, we would need to see p-values of < 0.00625 to believe that a predictor is associated with the compressive strength of concrete. **Cement, Blast.Furnace.Slag, and Age** are the only predictors that are associated with compressive strength afterwards.

## 3.

An extra kg of Cement in a $m^3$ mixture would increase the mean compressive strength of concrete by 0.117578.  
An extra day of waiting after pouring the concrete would increase the mean compressive strength of concrete by 0.127542.

## 4.

```{r}
fit.sw.bic = step(fit, k = log(length(concrete$Strength)))
summary(fit.sw.bic)
```

After pruning, the final regression equation is:
$$\widehat{Strength_i} = 40.23 + 0.11Cement_i+ 0.09Blast.Furnace.Slag_i + 0.08Fly.Ash_i - 0.29Water_i + 0.13Age_i$$

## 5.

### a.

```{r}
exampleData <- data.frame(Cement = 491,
                          Blast.Furnace.Slag = 26,
                          Fly.Ash=26,
                          Water = 210,
                          Superplasticizer = 3.9,
                          Coarse.Aggregate = 882,
                          Fine.Aggregate = 699,
                          Age = 28)

pred <- predict(fit.sw.bic, newdata=exampleData, interval="confidence", level=0.95)
pred
```

The predicted mean compressive strength for this mix is `r pred[1]`. The lower confidence interval is `r pred[2]` and the upper confidence interval is `r pred[3]`.

### b.

As the mix of concrete that the engineer is currently using has a greater mean compressive strength than the predicted mean compressive strength of the new concrete mix, the model suggests that the new mix is worse than the current mix. 

# Question 2.

## 1.
```{r}

```

